Formal verification of train route planning

### 2025-06-27 paper/main.tex
premiere version complete de la composition

So, we suppose we are given safe step states $S_1, \ldots, S_k$ corresponding to $X_1, \ldots, X_k$. To compute their composition $S_1 \cdot S_2 \cdots S_k$, we proceed by induction on $k$. For $k = 1$, we return $S_1$. For $k>0$, let $S = S_1 \cdots S_{k-1}  = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma, \langle E, W, T\rangle, \switches, \signals\rangle \rangle$. We set $S \cdot S_k = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma', \langle E', W, T\rangle, \switches', \signals\rangle \rangle$ as follows.

First,  for any train $\tid{i} \in \trains$, let $P_0 = S.\Gamma[\tid{i}].P$ and $P_k = S_k.\Gamma[\tid{i}].P$, then set $\Gamma'[\tid{i}] = \Gamma[\tid{i}][ P \leftarrow P_0 \cdot P_k]$.

Second, we proceed by updates of $E$ and $\sigma$. At first, we set $E' = E$ and $\sigma' = \sigma$. Now, suppose that $\sigma_k$ is the network configuration coming with $X_k$. Let $\sigma_0 = \{ \sid{i} \mid \sigma(\sid{i}) \neq  \sigma_k(\sid{i})\}$. Given  $\sid{i} \in \turnouts$, suppose there is a step $j < k$, a train $\tid{m}$ such that $\sid{i} \in \throughS(x_n, d, \sigma_j, x_{n+1})$ with $\vec{x}$ being the route of train $\tid{m}$ at step $X_j$ and $\sigma_j$ the network configuration associated to $j$. Then, we keep the triple  $\langle j, \tid{m}, n\rangle$ with the highest value $j$. If there is no such triple, set $\sigma' := \sigma'[ \sid{i} \leftarrow \sigma_k(\sid{i})]$.  Otherwise, set $E' := E'[ ....]$.

>>>>

So, we suppose we are given safe step states $S_1, \ldots, S_k$ corresponding to $X_1, \ldots, X_k$, network configuration $\sigma_1, \ldots, \sigma_k$ and direction maps $d_1, \ldots, d_k$. To compute their composition $S_1 \cdot S_2 \cdots S_k$, we proceed by induction on $k$. For $k = 1$, we return $S_1$. For $k>0$, let $S = S_1 \cdots S_{k-1}  = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma, \langle E, W, T\rangle, \switches, \signals\rangle \rangle$. We set $S \cdot S_k = \langle \guardG, \bufTrain, \bufSig, \langle \Gamma', \langle E', W, T\rangle, \switches', \signals\rangle \rangle$ as follows.

First,  for any train $\tid{i} \in \trains$, we append the orders of the train in $S_k$ to the corresponding ones in $S$. %let $P_0 = S.\Gamma[\tid{i}].P$ and $P_k = S_k.\Gamma[\tid{i}].P$, then set $\Gamma'[\tid{i}] = \Gamma[\tid{i}][ P \leftarrow P_0 \cdot P_k]$. 
Second, we proceed by updates of $E$ and $\sigma$. At first, we set $E' = E$ and $\sigma' = \sigma$. We begin to append to $E'$, for any train, all the handlers within $S_k$.

Consider some turnout $\sid{i}$ such that $\sid{i} \in \throughS(x_n, d, \sigma_k, x_{n+1})$ with $\vec{x}$ the route corresponding to train $\tid{i}$ in $X_k$. In other words, the turnout is crossed by train $\tid{i}$ at step $n$ of $\vec{x}$ within $X_k$. Suppose there is a step $j < k$, a train $\tid{m}$ such that $\sid{i} \in \throughS(y_m, d', \sigma_j, y_{m+1})$ with $\vec{y}$ being the route of train $\tid{m}$ at step $X_j$. If $\sigma_j(\sid{i}) = \sigma_k(\sid{i})$, we skip this turnout. Otherwise, if $\tid{i} = \tid{m}$, call $p$ the position within $E'$ corresponding to block $x_{n+1}$, we append in $E'$ at that position the order $\auth$. Otherwise, call $p$ the position in the map $E'$ corresponding to the atoms related to the block $y_{m+1}$. We append to that position the order $\turn{\sid{i}}{d_k(\sid{i})})$. In other words, we append to the last step involving the turnout the order for the new position.

Consider now a block within a route of $X_k$: $x_{n-1}^\tid{i}, x_n^\tid{i}, x_{n+1}^\tid{i}$. In case there is a route $y^\tid{m}$ within  $X_j$ such that\footnote{Again, we take the one with maximal value $j$ if there is a choice.} $y_{p-1}^\tid{m}, y_p^\tid{m}, y_{p+1}^\tid{m}$ with $x_n = y_p$ and $\tid{i} \neq \tid{m}$, we need to set a guard on the common resource $x_n = y_p$ that will be crossed by train $\tid{m}$ before train $\tid{i}$ will do it. We say that there is a conflict on $x_n$. We continue the process on $y_{p-1}^\tid{m}, y_p^\tid{m}, y_{p+1}^\tid{m}$, finding some subsequent conflicts. We do so up to the initial position. Call $n$ the number of conflicts met all along that process. Now, call $p$ the position within $E'$ corresponding to the position $x_{n-1}^\tid{i}$ and $q$ the position corresponding to $y_{p+1}^\tid{m}$. We append in $E'$ to $p$ the order $\wait{x_n}{n}$ and to $q$ the order $\incr{x_n}$.
